---
title: "lab 2 kc housing"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
real_estate<-read.csv("kc_house_data 2.csv")
real_estate$price_per_sqft <- real_estate$price/real_estate$sqft_living
real_estate$top_zip <- as.integer(real_estate$zipcode == 98004)
real_estate$top_5_zip <- as.integer(real_estate$zipcode == 98004 | real_estate$zipcode == 98005| real_estate$zipcode == 98052 | real_estate$zipcode == 98121 | real_estate$zipcode == 98109)
real_estate$top_10_zip <- as.integer(real_estate$zipcode == 98004 | real_estate$zipcode == 98005| real_estate$zipcode == 98052 | real_estate$zipcode == 98121 | real_estate$zipcode == 98109 |real_estate$zipcode == 98007 | real_estate$zipcode == 98122 | real_estate$zipcode == 98029 | real_estate$zipcode == 98102 | real_estate$zipcode == 98103)
real_estate$top_15_zip <- as.integer(real_estate$zipcode == 98004 | real_estate$zipcode == 98005| real_estate$zipcode == 98052 | real_estate$zipcode == 98121 | real_estate$zipcode == 98109 |real_estate$zipcode == 98007 | real_estate$zipcode == 98122 | real_estate$zipcode == 98029 | real_estate$zipcode == 98102 | real_estate$zipcode == 98103 | real_estate$zipcode == 98006 | real_estate$zipcode == 98033 | real_estate$zipcode == 98074 | real_estate$zipcode == 98040 | real_estate$zipcode == 98101)
real_estate$top_20_zip <- as.integer(real_estate$zipcode == 98004 | real_estate$zipcode == 98005| real_estate$zipcode == 98052 | real_estate$zipcode == 98121 | real_estate$zipcode == 98109 |real_estate$zipcode == 98007 | real_estate$zipcode == 98122 | real_estate$zipcode == 98029 | real_estate$zipcode == 98102 | real_estate$zipcode == 98103 |  real_estate$zipcode == 98006 | real_estate$zipcode == 98033 | real_estate$zipcode == 98074 | real_estate$zipcode == 98040 | real_estate$zipcode == 98101 | real_estate$zipcode == 98115 | real_estate$zipcode == 98112 | real_estate$zipcode == 98107 | real_estate$zipcode == 98119 | real_estate$zipcode == 98105)
```


```{r}
#plot of relationship between price per square ft and condition
y<-real_estate$price_per_sqft
x1<-real_estate$condition
plot(x1,y,main="Relationship between price per square ft and condition of housng",
     xlab="condition", ylab="Price per square ft of the house")
```

The plot shows the condition ranked 3 to have a high effect on 
the price per square ft of the house



```{r}
#Regression Modeling
#How does the condition of a house affect the home sale price?
model1<-lm(real_estate$price_per_sqft~real_estate$condition)
summary(model1)
```



From the regression model,the p-value is 2.2e-16 which is less 0.05 
unlikely we will observe a relationship between the predictor (conditions) and response (price) variables due to chance implying the test is significant and further we reject the true null 
hypothesis and conclude that the Conditions of a house have an effect on the home sale price per square ft.

It takes an average house in our dataset 204.634 conditions for the prices per sqr feet to be affected
for every 1 increase in the condition of a house, the Sale price per square feet of the house goes up by 17.452.
The adjusted r squared is 0.01061 which implies roughly 0.10% of the variance found in the response variable (price per square feet) can be explained by the predictor variable (condition)

The standard error shows the required condition  for the price per square ft of the house to be affected can vary by  109.4



```{r}
##Exploratory data analysis
#plot of relationship between prices and bedrooms
x2<-real_estate$bedrooms
y<-real_estate$price_per_sqft
plot(x2,y,main="Relationship between price per sqr ft and number of bedrooms",
     xlab="bedrooms", ylab="Price per square ft of the house")
```

The plot is skewwd to the left with most ratings centered between 1 and 10 bedrooms

```{r}
#plot of relationship between price per square ft and size of living room
x3<-real_estate$sqft_living
plot(x3,y,main="Relationship between price per square ft and sqft_living",
     xlab="sqft_living", ylab="Price per square ft of the house")
```

There is a slight skew to the left with most ratings centered between 0 and 6000



```{r}
#plot of relationship between price per square ft and square above the ground
x4<-real_estate$sqft_above
plot(x4,y,main="Relationship between price per square ft and sqft_above",
     xlab="sqft_above", ylab="Price per square ft of the house")
```

There is a slight skew to the left with most ratings centered between 2000 and 4000

```{r}
#plot of relationship between price per square ft and floors
x5<-real_estate$floors
plot(x5,y,main="Relationship between price per square ft and floors",
     xlab="floors", ylab="Price per square ft of the house")
```


There is a slight skew to the left with most ratings centered between 1 and 2.0 floors



```{r}
#plot of relationship between price per square ft and  grade
x6<-real_estate$grade
plot(x6,y,main="Relationship between price per square ft and grade",
     xlab="grade", ylab="Price per square ft of the house")
```

There is a slight skew to the right with most ratings centered between grade 5 and 12



```{r}
#plot of relationship between price per square ft and top 20 zip
x7<-real_estate$top_20_zip
plot(x7,y,main="Relationship between price per square ft and top 20 zip",
     xlab="top 20 zip code", ylab="Price per square ft of the house")
```

There is a slight skew to the left with most ratings symmetric between 0.0 and 1.0



```{r}
#plot of relationship between price per square ft and number of bathrooms
x8<-real_estate$bathrooms
plot(x8,y,main="Relationship between price per square ft and bathrooms",
     xlab="bathrooms", ylab="Price per square ft of the house")
```

There is a slight skew to the left with most ratings centered between 1 and 4



```{r}
##Asuumptions
model2<-lm(price_per_sqft~condition+bedrooms+bathrooms+top_20_zip+grade+floors+sqft_above
           +sqft_living,data = real_estate)
summary(model2)
```


From the regression model,the p-value is 2.2e-16 which is less 0.05 
unlikely we will observe a relationship between the predictors and 
response  variables due to chance



```{r}
#Diagonistics
par(mfrow = c(2, 2))
plot(model2)

##IID Assumption
plot(model2, 1)
```

The residual plot  shows no fitted pattern between the residuals 
and the fitted values hence IID


