---
title: "NBA Bayesian Analysis"
author: "Jana"
date: "2024-10-01"
output:
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Load necessary packages
library(ggplot2)
```


```{r}
#load data
NBAclutchFT <- read.csv("C:/Users/user/Desktop/R projects/Preply/2024 Fall/Jana/data/NBAclutchFT.csv")
str(NBAclutchFT)
```

**(a) Describe your model including the likelihood and prior.**

The dataset contains data on the number of successful "clutch" free throws made by nine NBA players and the number of attempts in these clutch situations. The aim is to analyze clutch performance for each player using Bayesian methods.

For each player, let *Yi* represent the number of successful clutch free throws, and *Ni* the total number of clutch free throw attempts. We model this as a binomial random variable:
   Yi|Î¸i ~ Binomial(Ni,Î¸i)

where:

Ni: the number of clutch free throw attempts.

Î¸i: the unknown probability of success (clutch FT percentage) for player i.

Yi: the number of successful clutch free throws.

**Likelihood**

For each player, the likelihood function is a binomial distribution.

**Prior**

The conjugate prior for a binomial likelihood is a Beta distribution:

theta i âˆ¼ Beta (alpha, beta)

Where ð›¼and Î² are the parameters of the prior, representing prior beliefs about the success rate of free throws. We can start with a non-informative or weakly informative prior such as:

ð›¼=1,ð›½=1(Uniform prior over [0,1])

This prior assumes no strong prior knowledge about the clutch free throw percentage.

**(b) Plot the posteriors of the clutch success probabilities.**

```{r}
# Given data
players <- c("Nikola Jokic", "Shai Gilgeous-Alexander", "Luka Doncic", 
             "Giannis Antetokounmpo", "Joel Embiid", "Jayson Tatum", 
             "Donovan Mitchell", "De'Aaron Fox", "Stephen Curry")
clutch_makes <- c(36, 35, 19, 39, 28, 33, 23, 14, 29)
clutch_attempts <- c(44, 39, 25, 53, 37, 40, 31, 19, 34)

# Set prior (weakly informative Beta(1, 1) prior)
alpha_prior <- 1
beta_prior <- 1

# Function to compute and plot posterior distributions
plot_posteriors <- function(makes, attempts, player) {
  # Posterior parameters
  alpha_post <- makes + alpha_prior
  beta_post <- attempts - makes + beta_prior
  
  # Generate theta values
  theta_vals <- seq(0, 1, length.out = 1000)
  posterior <- dbeta(theta_vals, alpha_post, beta_post)
  
  # Plot posterior
  df <- data.frame(theta = theta_vals, posterior = posterior)
  ggplot(df, aes(x = theta, y = posterior)) +
    geom_line(color = "blue") +
    labs(title = paste("Posterior of clutch FT% for", player),
         x = expression(theta), y = "Density")
}

# Plot posteriors for each player
for (i in 1:length(players)) {
  print(plot_posteriors(clutch_makes[i], clutch_attempts[i], players[i]))
}
```


**(c) Summarize the posteriors in a table.**

```{r}
# Function to compute posterior summary
posterior_summary <- function(makes, attempts) {
  alpha_post <- makes + alpha_prior
  beta_post <- attempts - makes + beta_prior
  
  mean_post <- alpha_post / (alpha_post + beta_post)
  lower_ci <- qbeta(0.025, alpha_post, beta_post)
  upper_ci <- qbeta(0.975, alpha_post, beta_post)
  
  return(c(mean_post, lower_ci, upper_ci))
}

# Create a summary table
summary_table <- data.frame(Player = players,
                            Mean = numeric(length(players)),
                            `Lower 95% CI` = numeric(length(players)),
                            `Upper 95% CI` = numeric(length(players)))

for (i in 1:length(players)) {
  summary_table[i, 2:4] <- posterior_summary(clutch_makes[i], clutch_attempts[i])
}

# Display summary table
print(summary_table)
```

The summary table presents the posterior mean and 95% credible intervals for the clutch free throw percentages of nine NBA players. Shai Gilgeous-Alexander has the highest estimated clutch free throw percentage with a mean of 87.8%, while De'Aaron Fox has the lowest at 71.4%. Nikola Jokic, Jayson Tatum, and Stephen Curry also show strong clutch performances, with means above 80%. The 95% credible intervals indicate the range within which the true clutch free throw percentage likely falls, with Stephen Curry's being particularly high (83.3%, CI: 69.7%-93.4%), showing consistent performance under pressure. Conversely, players like Donovan Mitchell and Luka Doncic have lower clutch percentages and wider credible intervals, suggesting more variability in their clutch performance.

**(d) Test the hypothesis that the clutch proportion is less than the overall proportion.**

```{r}
# Overall FT percentages (example values; replace with actual data)
overall_ft_pct <- c(0.817, 0.874, 0.786, 0.657, 0.847, 0.833, 0.836, 0.704, 0.856)

# Function to compute P(Î¸ < overall FT%)
posterior_prob <- function(makes, attempts, overall_ft) {
  alpha_post <- makes + alpha_prior
  beta_post <- attempts - makes + beta_prior
  
  prob <- pbeta(overall_ft, alpha_post, beta_post)
  return(prob)
}

# Compute posterior probabilities
posterior_probs <- numeric(length(players))
for (i in 1:length(players)) {
  posterior_probs[i] <- posterior_prob(clutch_makes[i], clutch_attempts[i], overall_ft_pct[i])
}

# Display posterior probabilities
posterior_probs
```

The hypothesis test results show strong evidence that Joel Embiid, Donovan Mitchell, Luka Doncic, Jayson Tatum, and Stephen Curry have clutch free throw percentages lower than their overall percentages, with posterior probabilities greater than 0.6, indicating moderate to strong support for this hypothesis. For Nikola Jokic, the evidence is weak, with a probability slightly above 0.5. However, for Shai Gilgeous-Alexander, Giannis Antetokounmpo, and De'Aaron Fox, the results suggest no significant difference between their clutch and overall free throw percentages, as their posterior probabilities are below or around 0.5.


**(e) Are the results sensitive to your prior?**


comparing three different priors:

Beta(1, 1) (our current uninformative prior).

Beta(5, 5) (a moderately informative prior assuming a mean success rate around 50%).

Beta(10, 10) (a stronger informative prior assuming a mean success rate around 50%).

*Compute Posterior Probabilities for Different Priors*

```{r}
# Given data for clutch FT
players <- c("Nikola Jokic", "Shai Gilgeous-Alexander", "Luka Doncic", 
             "Giannis Antetokounmpo", "Joel Embiid", "Jayson Tatum", 
             "Donovan Mitchell", "De'Aaron Fox", "Stephen Curry")
clutch_makes <- c(36, 35, 19, 39, 28, 33, 23, 14, 29)
clutch_attempts <- c(44, 39, 25, 53, 37, 40, 31, 19, 34)

# Overall FT percentages
overall_ft_pct <- c(0.817, 0.874, 0.786, 0.657, 0.847, 0.833, 0.836, 0.704, 0.856)

# Function to compute posterior probability with different priors
posterior_prob_with_prior <- function(makes, attempts, overall_ft, alpha_prior, beta_prior) {
  alpha_post <- makes + alpha_prior
  beta_post <- attempts - makes + beta_prior
  prob <- pbeta(overall_ft, alpha_post, beta_post)
  return(prob)
}

# Different priors to test
priors <- list(c(1, 1), c(5, 5), c(10, 10))
prior_names <- c("Beta(1,1)", "Beta(5,5)", "Beta(10,10)")

# Create an empty matrix to store posterior probabilities
posterior_probabilities <- matrix(nrow = length(players), ncol = length(priors))

# Loop through each prior and compute the posterior probabilities for each player
for (i in 1:length(players)) {
  for (j in 1:length(priors)) {
    prior <- priors[[j]]
    posterior_probabilities[i, j] <- posterior_prob_with_prior(clutch_makes[i], clutch_attempts[i], overall_ft_pct[i], prior[1], prior[2])
  }
}

# Convert the matrix to a data frame for easy reading
posterior_probs_df <- data.frame(
  Player = players,
  `Beta(1,1)` = posterior_probabilities[, 1],
  `Beta(5,5)` = posterior_probabilities[, 2],
  `Beta(10,10)` = posterior_probabilities[, 3]
)

# Display the results
print(posterior_probs_df)

```

The results show that the posterior probabilities are somewhat sensitive to the choice of prior. For players like Nikola Jokic, Shai Gilgeous-Alexander, Luka Doncic, and De'Aaron Fox, the probabilities that their clutch free throw percentages are lower than their overall percentages significantly increase as the prior becomes more informative. This suggests that the data alone isn't strong enough to dominate the prior, making the results sensitive to prior assumptions. However, for players like Joel Embiid and Donovan Mitchell, the results remain consistently high across all priors, indicating robust evidence that their clutch free throw percentages are lower, regardless of the prior used. Overall, stronger priors increase the probability of lower clutch performance for most players.


**2. Say that Y|Î¸ ~ Binomial(N,Î¸) and Z|Î¸~Binomial(M,Î¸), and that Y and Z are independent given Î¸. Identify a conjugate prior for Î¸ and find the corresponding posterior distribution.**

```{r}
# Set parameters for binomial observations
N <- 50  # Number of trials for Y
Y <- 30  # Number of successes in Y
M <- 40  # Number of trials for Z
Z <- 20  # Number of successes in Z

# Set prior parameters (Beta prior)
alpha_prior <- 2  # Prior successes
beta_prior <- 2   # Prior failures

# Compute posterior parameters
alpha_post <- alpha_prior + Y + Z
beta_post <- beta_prior + (N - Y) + (M - Z)

# Generate theta values between 0 and 1
theta_vals <- seq(0, 1, length.out = 1000)

# Compute the prior and posterior densities
prior_density <- dbeta(theta_vals, alpha_prior, beta_prior)
posterior_density <- dbeta(theta_vals, alpha_post, beta_post)

df <- data.frame(theta = theta_vals, 
                 Prior = prior_density, 
                 Posterior = posterior_density)

ggplot(df, aes(x = theta)) +
  geom_line(aes(y = Prior), color = "red", linetype = "dashed") +
  geom_line(aes(y = Posterior), color = "blue") +
  labs(title = "Prior vs Posterior Distributions for Î¸",
       x = expression(theta), 
       y = "Density") +
  theme_minimal()

# Posterior mean and credible interval
mean_post <- alpha_post / (alpha_post + beta_post)
median_post <- qbeta(0.5, alpha_post, beta_post)
lower_ci <- qbeta(0.025, alpha_post, beta_post)
upper_ci <- qbeta(0.975, alpha_post, beta_post)

# Display summary
posterior_summary <- data.frame(
  Mean = mean_post,
  Median = median_post,
  `Lower 95% CI` = lower_ci,
  `Upper 95% CI` = upper_ci
)

print(posterior_summary)


```

